Caring 15-minute neighbourhoods

### 2024-10-10 paper/Manuscript.tex
Edits before submission

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

@@

Though the term 15-Minute City is relatively new, neighbourhood planning
practice aiming to foster community and local travel to amenities is
not. As recent examples, planned neighbourhoods dominated post-WWII
built urban form: the Neighbourhood Unit Concept in the western world
\citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
``mikrorayons'' neighbourhoods in the Soviet Union
\citep{kissfazekasCircleParadigms15minute2022}. Along different
dimensions, these planning forms have been extensively critiqued
\citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
planning approaches such as the 15-Minute City concept are being
re-embraced by some, but from bottom-up perspective using co-creation
tools and meaningful resident participation to achieve equitable and
justice neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}.
How to equitably prescribe urban form that `enables' contact with social
opportunities (instead of `engendering it' from the top-down) is still
under debate. Questions remain, including: what destinations really
matter?, what mode and travel time threshold is most appropriate?, and
who benefits? For instance, these dimensions are under discussion among
proximity-based planners who use accessibility-based tools
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.

From another perspective, feminist geographers, urban planners and other
researchers agree that urban structure influences travel, but city
planning falls short in planning for a neutral identity. In contrast,
cities ought to be planned according to the multiple identities of the
inhabitants
\citep{vacchelli_towards_2018, urbandevelopmentviennaGenderMainstreamingUrban2013}.
Researchers have demonstrated how mobility patterns differ according to
gender
\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990}.
Gendered social norms and cultural factors play a role in how, where,
with whom, what for, and how far we move. Women have higher activities
related to care \citep{ilo_care_2018}, no matter at what life stage they
are \citep{garciaromanGenderDifferencesTime2022}. For all these reasons
the spatial organization of activities has a different impact according
to gender, the 15-Minute City concept should not be gender blind.

This work builds a theoretical bridge between the 15-minute city and
mobility of care concepts to answer questions about what destinations
matter and who benefits from proximity planning. The Mobility of Care
term was coined in
\citet{sanchezdemadariagaMobilityCareIntroducing2013}; it refers to all
travel required to sustain the needs of a household, such as grocery
shopping, escorting children, travelling to health appointments, and
running errands. While decades of research have examined types of
household-serving trips (e.g., shopping trips, escorting to school),
\citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
consider them all as one category. In doing so, Mobility of Care is
shown to be a significant proportion of daily travel, characterizing the
trip purpose of approximately 30\% of adults' trips
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
We argue Mobility of Care is especially relevant to the 15-Minute City,
as care trips are necessary for \emph{all} people and they most often
occur at the local level. Moreover, care trips take on tremendous
gendered and socio-economic importance, as women, and especially those
from lower income households, complete the majority of care trips
\citep{ravensbergen2023exploratory}. Despite care travel's importance in
advancing gender equality -and equity more broadly-, mobility of care
and access to care destinations have been under-examined relative to
employment destinations. This is especially pertinent in accessibility
research that has largely focused on employment points of interest e.g.,
\citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.

The objectives of this research are two-fold. First, we aim to
theoretically bridge Mobility of Care with the 15-Minute City concept to
re-imagine what local amenities matter and to define an associated
accessibility and diversity of accessibility measures. Second, we take a
data-driven approach to classify areas into `caring 15-minute
neighbourhoods' typologies and examine associated residential profiles
in an empirical examination of the city of Hamilton, Canada.

@@

In the last decades, the need for more sustainable, healthier, and
livable cities have received significant attention. Planners and
decision-makers have proposed alternatives to mobility-based
neighbourhood planning principles to ones that center proximity to urban
functions \citep{pozoukidou15MinuteCityDecomposing2021}. In this
context, the 15-Minute City is now in the public spotlight
\citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
was introduced in \citet{moreno_introducing_2021} along with four
dimensions: density (in terms of people per km\(^2\)), diversity
(including mixed-land use and diversity of people), the temporal and
spatial proximity to essential services, and digitalisation. However,
the framework presented has been criticized within academic and planning
circles \citep{guzmanProximityEnoughCritical2024}. Critiques highlight
how the concept falls short in addressing the pre-existing structural
forces and individual characteristics that drive inequalities,
influencing who benefits or could benefit the most from such a paradigm
\citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
15-Minute City for all people is an aspirational ideal, but how will
existing mobility and accessibility inequalities be addressed. Without
directed and context-specific solutions, this popular concept risks
becoming an empty city branding exercise.
\citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.

Reflecting the flexible and aspirational presentation of the 15-Minute
City; cities that have adopted the concept have done so using a diverse
range of definitions and tools along with indistinct universal
approaches. A trail blazer was Portland, US in the Portland Plan
\citep{city_of_portland_20minute_2010} of April 2012. Adopted before the
`15-Minute City' concept of today, it aims to foster an inclusive city
development based on prosperity, education, health, and equity over a 30
year horizon. The plan aims to promote neighbourhood self-sufficiency
and connectivity to city centers and centres of employment. The progress
report describes a high-level focus on equitable service delivery to all
residents with equity concerning topics related to racial equity and
people with disabilities \citep{portland_government_portland_2017},
similarly taking on an ``all populations'' approach. Later, other cities
adopted proximity-based goals using similar neutral approaches. For
example, the 15-Minute City plan announced during the Paris mayor Anne
Hidalgo's 2020 re-election campaign, which helped popularize the concept
in today's public discourse, emphasized six key social functions that
should be easily accessible from any location
\citep{ville_de_paris_paris_2022}. These locations included: housing,
work, health care, groceries, education and leisure. The 15-Minute City
concept inspired language in the agendas of other cities in the Western
world such as Ottawa, in the Canadian context, who also adopted a
15-Minutes approach in their recent Official Plan
\citep{ville_dottawa_quartier_2021}.
\citet{teixeiraClassifying15minuteCities2024}'s global review finds that
the 15-Minute City concept is in early phases of implementation around
the world and the diverse range of definitions, strategies, and
instruments present a significant knowledge and implementation gap. In
other words: the 15-Minute City is aspirational but how do we get there
and who will benefit?

The past can shed light on cautionary tales. While the 15-Minute City
concept is new, neighbourhood planning to improve society outcomes is
not: in fact the literature has drawn parallels from the 15-Minute City
to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s
\citep{kissfazekasCircleParadigms15minute2022}. The NUC is a
socio-spatial normative scheme widely adopted by government agencies
(and the real-estate community) in the Western world after the Second
World War \citep{talenSocialSciencePlanned2017, solow1969concept}.
Pairing well with the objectives of planning agencies at the time,
Perry's NUC would allow for efficient mass-building of cellular units
that prioritized the perceived functional needs of women and children:
each unit providing proximate access to an elementary school and
supporting community facilities, shopping, parks and housing
\citep{talenSocialSciencePlanned2017, brodyNeighbourhoodUnitConcept2013}.
The NUC primarily prioritized local service provision, though Perry had
confidence in good design's contribution to `neighbourhood spirit'
\citep{hall2014cities}. From the NUC's conception until the end of the
1960s, planners' aspirational attempts to prescribe social meaning to
the neighbourhood's physical form had been criticized to near extinction
\citep{talenSocialSciencePlanned2017}. By social scientists, the
overestimation on the impact of built environment on social life was
critiqued. By other planners, the inability to agree on the specificity
of the neighbourhood (i.e., population size and the type, quality, and
quantity of amenity) as well as how neighbourhood units are connected to
other neighbourhoods and the rest of the region. In response to these
criticisms, neighbourhood planning proponents redefined deterministic
terminology; their prescriptive physical form will `enable' social
contact with opportunities rather than `engendering' it.

The redefined bottom-up approach to community and neighbourhood planning
has been adopted by New Urbanists, from which the 15-Minute City stems
\citep{kissfazekasCircleParadigms15minute2022}; however, the
effectiveness of these bottom-up processes remain to be seen due to the
contemporary nature of the concept. Though, related research can be
examined. In recent years, the question of \emph{how can physical form
be planned to enable an improved quality of life, for whom, and with
what outcomes} has occupied the urban planning research agenda. For
instance, the examination of low-income households access to
transportation and their likelihood of gaining employment
\citep{blumenbergDriveWorkRelationship2017, bastiaanssenDoesBetterJob2022}
or the relationship between children's access to public transit and
participation in after-school activities
\citep{palmRolePublicTransit2020}. A new wave of researchers and
practitioners focused on local and context-specific relationships with
the proximity to destinations have also emerged
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022}.
As reviewed in the city plans that have adopted 15-Minute City
approaches, the NUC's criticisms, and recent urban planning research,
the question of \emph{how to enable improved quality of life through
urban built environment} has shown to be highly context-sensitive, with
the need for further investigation.

@@

In examining how to enable improved quality of life through urban built
environment, accessibility measures have become an increasingly popular
tool. These measures are a way to quantify the ease of reaching
destinations from a given point in space and have been used to examine
urban areas through just and sustainable city planning agendas
\citep{vale2023accessibility, handyAccessibilityIdeaWhose2020}. The
15-Minute City is an amenity-provision neighbourhood planning concept
aimed at enabling the creation of urban environments that enhance life
quality, making it well-suited for analysis using accessibility
methodologies \citep{guzmanProximityEnoughCritical2024}. Recent works
have applied accessibility measures to investigate the 15-Minute City
across different geographic scopes. For instance, in Naples, Italy
\citep{gaglioneUrbanAccessibility15minute2022}, Barcelona, Spain
\citep{graellsCityCitiesMeasuring2021}, Vancouver, Canada
\citep{hosford15minuteCityReach2022}, and in urban areas across Europe
\citep{vale2023accessibility}. The ``cumulative opportunity'' measure
has been applied in many 15-Minute City examinations. This measure
quantifies how many destinations can be reached from a point in space
within a given travel time threshold, pairing well with examinations
that have a normative travel time threshold, like the 15-Minute City.
Furthermore, the cumulative opportunity measure is widely appreciated
for its intuitive computation and popularity among practitioners
\citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
However, accessibility measures other than the cumulative opportunity
have also been applied, reflecting the diversity of measures in the
literature \citep{guzmanProximityEnoughCritical2024}.

Another method that compliments the assessment of the 15-Minute City
regarding the diversity of opportunity types is entropy. The entropy
measure, based on the Shannon-Wiener index of species diversity,
expresses relative evenness within a sequence
\citep{whittaker1972evolution}. In urban studies and planning
literature, the flexible concept of entropy has been used to
characterize land-use mix diversity
\citep{frankLinkingObjectivelyMeasured2005, ewingTravelBuiltEnvironment2010a},
particularly in the context of non-work trips
\citep{cerveroTravelDemand3Ds1997}, walking as a mode
\citep{luUrbanDensityDiversity2017, mavoaIdentifyingAppropriateLanduse2018},
and suburban sprawl \citep{randallGISBasedLand2015} as relevant to this
work. It has also been used to understand mobility behaviour
\citep{mcbride2020exploration, montero2023applying, montero2023role}.
However, work incorporating entropy indices alongside accessibility
analysis are not exceptionally common. There are examples of their use
as parameters within accessibility scores for employment opportunities
\citep{chengMeasuringUrbanJob2013, daiIncorporatingJobDiversity2018}
and, more recently, to describe the diversity in transit facilities
\citep{yinIncorporatingFacilityDiversity2024}. Within the
conceptualisation of the 15-Minute City, diversity of land-use is a key
feature. In this way, the entropy measure theoretically complements the
15-Minute City.

Classification algorithms also show promise as a tool in the assessment
of the 15-Minute City, as they have been useful in identifying mobility
and spatial typologies within the transportation planning literature.
Often, these algorithms take the form of machine learning approaches:
for instance, \citet{burkeGeospatialAnalysisFramework2022} uses
spatially constrained multivariate clustering, to develop urban form
typologies related to the 15-Minute City. The k-means algorithm is
another popular method: \citet{ganUnderstandingUrbanMobility2020}
applies it to analyse travel behaviour and classify metro stations based
on mobility patterns. Similarly,
\citep{montero2023role, montero2023applying} cluster travel behavior
profile typologies representing different population segments and place
of residence in the Barcelona Metropolitan Area. Combing classification
techniques is also common: for instance,
\citet{delmelleDifferentiatingPathwaysNeighborhood2017} uses
self-organizing maps (SOM) algorithm to group U.S. neighbourhoods by
minimizing dissimilarity in attributes before applying k-means for
further classification. Similarly, \citet{victorianoTimeSpaceMoney2020}
uses SOM to classify individuals' travel patterns according to the
dissimilarity in mobility attributes and uses the Decision Tree
algorithm to partition the data into interpretable classifications.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

@@

The term 15-Minute City is relatively new but the concept is not, as it
is simply a normative cumulative-opportunities accessibility measure
\citep{paez_measuring_2012}. Its newfound fame under the moniker of
chronourbanism comes to complement long-standing efforts to foster
community and local travel to amenities in neighbourhood planning
practice. As recent examples, planned neighbourhoods dominated post-WWII
built urban form: the Neighbourhood Unit Concept in the western world
\citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
``mikrorayons'' neighbourhoods in the Soviet Union
\citep{kissfazekasCircleParadigms15minute2022}. Along different
dimensions, these planning forms have been extensively critiqued
\citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
planning approaches such as the 15-Minute City concept offer an
opportunity for bottom-up approaches that leverage co-creation tools and
meaningful resident participation to achieve equitable and just
neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}. How to
prescribe equitable urban forms that `enable' contact with social
opportunities (instead of `engendering it' from the top-down) is still a
matter for debate. Questions remain, including: what destinations really
matter? What mode and travel time threshold is most appropriate? And who
benefits? To be certain, these dimensions are under discussion among
proximity-based planners who use accessibility-based tools
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.

From another perspective, feminist geographers, urban planners and other
researchers agree that urban structure influences travel, but city
planning falls short in planning for a neutral identity. In contrast,
cities ought to be planned according to the multiple identities of the
inhabitants
\citep{vacchelli_towards_2018, urbandevelopmentviennaGenderMainstreamingUrban2013}.
Researchers have demonstrated how mobility patterns differ according to
gender
\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990, neto_gender_2015}.
The causes for this are varied, and related to gendered social norms and
cultural factors that play a role in how, where, with whom, what for,
and how far we move. Of note is that women tend to travel more
frequently and for shorter distance
\citep{roorda_triap_2010, morency_distance_2011} and have more
activities related to care \citep{ilo_care_2018} no matter at what life
stage they are \citep{garciaromanGenderDifferencesTime2022}. Combined,
these findings indicate that the spatial organization of activities has
a different impact according to gender, and therefore the 15-Minute City
concept should not be gender blind.

This work builds a theoretical bridge between the 15-Minute City and
mobility of care concepts to answer questions about what destinations
matter and who benefits from proximity planning. The Mobility of Care
term was coined in
\citet{sanchezdemadariagaMobilityCareIntroducing2013}; it refers to all
travel required to sustain the needs of a household, such as grocery
shopping, escorting children, travelling to health appointments, and
running errands. While decades of research have examined types of
household-serving trips (e.g., shopping trips, escorting to school),
\citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
consider them all as one category. In doing so, care-related mobility
can be shown to comprise a significant proportion of daily travel,
characterizing approximately 30\% of adults' daily trips
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
We argue that Mobility of Care is especially relevant to the 15-Minute
City, as care trips are necessary for \emph{all} people and they most
often occur at the local level. Moreover, care trips are of particular
gendered and socio-economic import, as women, and especially those in
lower income households, complete the majority of care trips
\citep{ravensbergen2023exploratory}. Despite care travel's importance in
advancing gender equality -and equity more broadly-, mobility of care
and access to care destinations have been under-examined relative to
employment destinations. This is especially pertinent in accessibility
research that has largely focused on employment points of interest e.g.,
\citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.

The objectives of this research are two-fold. First, we aim to
theoretically bridge Mobility of Care with the 15-Minute City concept to
re-imagine what local amenities matter from the perspective of care and
to define an associated accessibility and diversity of accessibility
measures. And secondly, we take a data-driven approach to classify areas
into `caring 15-minute neighbourhoods' typologies and examine associated
residential profiles in an empirical examination of the city of
Hamilton, Canada.

@@

In the last decades, the need for more sustainable, healthier, and
livable cities has become a prominent concern. Planners and
decision-makers have proposed a shift to neighbourhood planning
principles centered on proximity to urban functions
\citep{pozoukidou15MinuteCityDecomposing2021}. In this context, the
15-Minute City is now in the public spotlight
\citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
was introduced in \citet{moreno_introducing_2021} along with four
dimensions: density (in terms of people per km\(^2\)), diversity
(including mixed-land use and diversity of people), the temporal and
spatial proximity to essential services, and digitalisation
\citep[q.v.][]{cerveroTravelDemand3Ds1997}. However, the framework
presented has been criticized within academic and planning circles
\citep[e.g.,][]{guzmanProximityEnoughCritical2024, mouratidisTimeChallenge15minute2024}
among other things for shortfalls in terms of addressing pre-existing
structural forces and individual characteristics that drive inequalities
and influence who benefits or could benefit the most from such an
approach
\citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
15-Minute City for all people is an aspirational goal, but does not
fully confront existing mobility and accessibility inequalities in need
of redress. Without directed and context-specific solutions, this
popular concept risks becoming an empty city branding exercise.
\citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.

Reflective of the flexible and aspirational presentation of the
15-Minute City, the concept has been adopted by cities using a diverse
range of definitions and tools along with indistinct universal
approaches. A trail blazer in this respect was Portland (US) in the
Portland Plan \citep{city_of_portland_20minute_2010} of April 2012.
Adopted before the `15-Minute City' concept of today, this plan aimed to
foster inclusive city development based on prosperity, education,
health, and equity over a 30 year horizon. Central to the plan was the
promotion of neighbourhood self-sufficiency and connectivity to city
centers and centres of employment. The progress report describes a
high-level focus on equitable service delivery to all residents with
equity concerning topics related to racial equity and people with
disabilities \citep{portland_government_portland_2017}, similarly taking
on an ``all populations'' approach. Subsequently, other cities adopted
proximity-based goals using similarly neutral approaches. For example,
the 15-Minute City plan announced during the re-election campaign of
Paris mayor Anne Hidalgo in 2020 emphasized six key social functions
that should be easily accessible from any location
\citep{ville_de_paris_paris_2022}. These locations included: housing,
work, health care, groceries, education and leisure. The 15-Minute City
concept inspired language in the agendas of other cities in the Western
world such as Ottawa, in the Canadian context, who also adopted a
15-Minutes approach in their recent Official Plan
\citep{ville_dottawa_quartier_2021}.
\citet{teixeiraClassifying15minuteCities2024}'s global review finds that
the 15-Minute City concept is in early phases of implementation around
the world and the diverse range of definitions, strategies, and
instruments present a significant knowledge and implementation gap. In
other words: the 15-Minute City is aspirational but how do we get there
and who will benefit?

The past can provide cautionary tales. While the 15-Minute City brand of
planning is new, neighbourhood planning to improve society outcomes is
not: in fact the literature has drawn parallels from the 15-Minute City
to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s
\citep{kissfazekasCircleParadigms15minute2022}. The NUC is a
socio-spatial normative scheme widely adopted by government agencies
(and the real-estate community) in the Western world after the Second
World War \citep{talenSocialSciencePlanned2017, solow1969concept}.
Pairing well with the objectives of planning agencies at the time,
Perry's NUC would allow for efficient mass-building of cellular units
that prioritized the perceived functional needs of women and children:
each unit providing proximate access to an elementary school and
supporting community facilities, shopping, parks and housing
\citep{talenSocialSciencePlanned2017, brodyNeighbourhoodUnitConcept2013}.
The NUC primarily prioritized local service provision, though Perry had
confidence in good design's contribution to `neighbourhood spirit'
\citep{hall2014cities}. By the end of the 1960s, planners' aspirational
attempts to prescribe social meaning to the neighbourhood's physical
form had been criticized to near extinction
\citep{talenSocialSciencePlanned2017}. A critique by social scientists
was an overestimation of the impact of built environment on social life.
Planners, on the other hand, could not reach consensus on the
specificity of the neighbourhood (i.e., population size and the type,
quality, and quantity of amenity) or how neighbourhood units connect
between them and the rest of the region. In response to these
criticisms, neighbourhood planning proponents redefined their
deterministic terminology; their prescriptive physical form would
`enable' social contact with opportunities rather than `engendering' it.

The redefined bottom-up approach to community and neighbourhood planning
was adopted in the 1980s by American New Urbanists
\citep{trudeau_new_2013}, from which the 15-Minute City eventually
stemmed \citep{kissfazekasCircleParadigms15minute2022}. However, the
bottom-up effectiveness of these ideas remains to be seen due to the
contemporary nature of the concept. Though, related research can be
examined. In recent years, the question of \emph{how can physical form
be planned to enable an improved quality of life, for whom, and with
what outcomes} has occupied the urban planning research agenda. For
instance, the examination of low-income households access to
transportation and their likelihood of gaining employment
\citep{blumenbergDriveWorkRelationship2017, bastiaanssenDoesBetterJob2022}
or the relationship between children's access to public transit and
participation in after-school activities
\citep{palmRolePublicTransit2020}. A new wave of researchers and
practitioners focused on local and context-specific relationships with
the proximity to destinations have also emerged
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022}.
As reviewed in the city plans that have adopted 15-Minute City
approaches, the NUC's criticisms, and recent urban planning research,
the question of \emph{how to enable improved quality of life through
urban built environment} is highly context-sensitive, prompting the need
for further investigation.

@@

In examining how to enable improved quality of life through urban built
environment, accessibility measures have become an increasingly popular
tool. These measures are a way to quantify the ease of reaching
destinations from a given point in space and have been used to examine
urban areas through just and sustainable city planning agendas
\citep{vale2023accessibility, handyAccessibilityIdeaWhose2020}. The
15-Minute City is an amenity-provision neighbourhood planning concept
aimed at enabling the creation of urban environments that enhance life
quality, making it well-suited for analysis using accessibility
methodologies \citep{guzmanProximityEnoughCritical2024}. Recent works
have applied accessibility measures to investigate the 15-Minute City
across different geographic scopes. For instance, in Naples, Italy
\citep{gaglioneUrbanAccessibility15minute2022}, Barcelona, Spain
\citep{graellsCityCitiesMeasuring2021}, Vancouver, Canada
\citep{hosford15minuteCityReach2022}, and in urban areas across Europe
\citep{vale2023accessibility}. The ``cumulative opportunity'' measure
has been applied in many 15-Minute City examinations. This measure
quantifies how many destinations can be reached from a point in space
within a given travel time threshold , pairing well with normative
examinations \citep[see][]{paez_measuring_2012} that use travel time
thresholds like x-minute cities {[}\citet{logan_xminute_2022}.
Furthermore, the cumulative opportunity measure is widely appreciated
for its intuitive computation and popularity among practitioners
\citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
However, accessibility measures other than the cumulative opportunity
have also been applied, reflecting the diversity of measures in the
literature \citep{guzmanProximityEnoughCritical2024}.

Another concept that complements the assessment of the 15-Minute City
regarding the diversity of opportunity types is entropy. The entropy
measure, based on the Shannon-Wiener index of species diversity,
expresses relative evenness within a sequence
\citep{whittaker1972evolution}. In urban studies and planning
literature, the concept of entropy has been used to characterize
land-use mix diversity
\citep{frankLinkingObjectivelyMeasured2005, ewingTravelBuiltEnvironment2010a, moniruzzaman_accessibility_2012}
including to understand mobility behaviour
\citep{mcbride2020exploration, montero2023applying, montero2023role}, in
the context of non-work trips \citep{cerveroTravelDemand3Ds1997},
walking
\citep{luUrbanDensityDiversity2017, mavoaIdentifyingAppropriateLanduse2018},
and suburban sprawl \citep{randallGISBasedLand2015}. However, entropy
indices are rare in accessibility analysis. There are examples of their
use as parameters within accessibility scores for employment
opportunities
\citep{chengMeasuringUrbanJob2013, daiIncorporatingJobDiversity2018}
and, more recently, to describe the diversity in transit facilities
\citep{yinIncorporatingFacilityDiversity2024}. Given how 15-Minute
Cities aim to provide access to numerous amenities, diversity of
land-use is a key feature. In this way, the entropy measure
theoretically complements the 15-Minute City.

Classification algorithms also show promise as a tool in the assessment
of the 15-Minute City, as they have been useful in identifying mobility
and spatial typologies within the transportation planning literature.
Often, these algorithms take the form of machine learning approaches: in
transportation, the use of Self-Organizing Maps (SOM) was pioneered by
Delmelle \citeyearpar{delmelle_spatial_2012} and has been used to group
U.S. neighbourhoods by minimizing dissimilarity in attributes
\citep{delmelleDifferentiatingPathwaysNeighborhood2017} and to classify
individuals' travel patterns according to the dissimilarity in mobility
attributes and uses the Decision Tree algorithm to partition the data
into interpretable classifications \citep{victorianoTimeSpaceMoney2020}.
Other approaches include the use of spatially constrained multivariate
clustering, to develop urban form typologies related to the 15-Minute
City \citep{burkeGeospatialAnalysisFramework2022} and the use of k-means
to analyse travel behaviour and classify metro stations based on
mobility patterns\citep{ganUnderstandingUrbanMobility2020}.
